TacticToe: Learning to Prove with Tactics

### 2018-08-17 tactictoe_jv.tex
theorem

In general, the list of arguments in such tactics respect the two following 
property:
\begin{itemize}
\item Additional theorems do not cause the tactic to fail and useless theorems 
are ignored. For \texttt{REWRITE\_TAC}, loops are generally avoided because 
most equations appears with only one orientation in the database of theorems.
\item The order of theorems in the list does not influence the final result and 
thus we can consider the list to be a set in our algorithm.
\end{itemize}

Since our predictor is not perfectly accurate and the presented tactics are 
robust, we select a list of theorems usually longer that the original one to 
instantiate these tactics.
Our default predictor for list of theorems selects 16 theorems from the 
database of theorems $\mathbb{D}_{\mathit{theorem}}$. It uses the produced list
$\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= [thm1,\ldots,thm16]$ to 
replace the
placeholders in $\hat{t}$. The instantiation of the first abstracted is thus
\texttt{REWRITE\_TAC [thm1,...,thm16]}.

\paragraph{Theorem}
Instantiating a single theorem is much harder than a list of theorems. Indeed,
the predictions needs to be very precise for the tactic to be effective.
For example, the tactic MPTAC only works if the theorem is an instance
of the precondition. The best option would be to have a train a theorem 
predictor that can adapt its predictions to the given tactic. However, this 
would require a stronger machine learning model than nearest neighbor algorithm.
Therefore, in order to mitigate the gap in the quality of the predictions, we 
propose to create a tactic that tries instantiations until one succeeds. (i.e 
by succeeding we mean produces a list of goal that does not contain the 
original goal)

@@

Side experiments with term argument predictions 
show that trying to instantiate the tactic with subterms 
from the goal improve the success rate by another 4 percent. 
However, in the evaluation presented in this paper, the 
abstraction/instantiation method is evaluated only for the type of theorem list.
Moreover, the current argument predictions are only affected by the given goal, 
and we believe it would be beneficial to have more flexible predictors that can 
adapt to their predictions to the abstracted tactic.

>>>>

In general, the list of arguments in such tactics respect the two following 
property:
\begin{itemize}
\item Additional theorems do not cause the tactic to fail and useless theorems 
are ignored. For \texttt{REWRITE\_TAC}, loops are generally avoided because 
most equations appears with only one orientation in the database of theorems.
\item The order of theorems in the list does not influence the final result and 
thus we can consider the list to be a set in our algorithm. This is 
particularly true for the first-order theorem prover \texttt{METIS\_TAC}.
\end{itemize}

Since our predictor is not perfectly accurate and the presented tactics are 
robust, we select a list of theorems usually longer that the original one to 
instantiate these tactics.
Our default predictor for list of theorems selects 16 theorems from the 
database of theorems $\mathbb{D}_{\mathit{theorem}}$. It uses the produced list
$\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= [thm1,\ldots,thm16]$ to 
replace the
placeholders in $\hat{t}$. Thus, the instantiation of the first construction 
would be \texttt{REWRITE\_TAC [thm1,...,thm16]}.

\paragraph{Theorem}
To instantiate a single theorem correctly is much harder than finding a 
superset of theorems. The following examples with placeholder \texttt{thm} 
illustrate this point:
  \texttt{MATCH\_MP\_TAC thm}, \texttt{INDUCT\_THEN thm ASSUME\_TAC}

Here, the two examples accepts only very specific kind of theorems: the theorem 
argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
and an induction principle should be given to \texttt{INDUCT\_THEN}.
The likelihood of finding such theorem is low, thus given the list of 
predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
[thm1,\ldots,thm16]$, we create 16 instantiated tactics and try them in the 
order given by the predictor, and we return the result of the first successful 
tactic and produces new goals. For readers familiar with \holfour tacticals, we 
give the single new tactic that can be created using this procedure on the 
first example:  
  \texttt{CHANGED\_TAC (MATCH\_MP\_TAC thm1) ORELSE ... ORELSE CHANGED\_TAC 
  (MATCH\_MP\_TAC thm16)}.

@@

Side experiments with term argument predictions 
show that trying to instantiate the tactic with subterms 
from the goal improve the success rate by another 4 percent. 
However, in the evaluation presented in this paper, the 
abstraction/instantiation method is evaluated only for the type of theorem list.
Moreover, the current argument predictions are only influenced by the input 
goal, 
and we believe that in the future it would be beneficial to have more flexible 
predictors that can adapt their predictions to each abstracted tactic.
